---
title: "Plastic waste management in Switzerland"
format:
  html: default
date: 06-06-2024
author:
  - name: Margaux Götschmann
    affiliation:
      - name: 'ETH Zürich'
table-of-contents: true
execute: 
  echo: true
  warning: false
abstract: ""
keywords:
  - plastic waste
  - waste management
editor: visual
bibliography: bibliography.bib
---

```{css, echo = FALSE}
.justify {
  text-align: justify !important
}
```
::: {.justify}
This project is part of an assignment for the course 'Research Beyond the Lab: Open Science and Research Methods for a Global Engineer' taught at ETH Zürich.

![[Source](https://www.ebp.ch/en/node/1225): Plastic Recycling in Switzerland \| EBP \| Swiss](plastic.jpeg)
:::
# Introduction
::: {.justify}
Switzerland faces significant challenges in recycling plastic waste. Each person in the country produces an average of 95 kg of plastic waste annually [@2023europe]. According to Patrick Semadeni, vice president of KUNSTOFF.swiss, Switzerland lacks the separation and recycling capacity to process all the collected plastic waste effectively [@turuban2024why]. This survey seeks to gain insights into the plastic waste management practices of individuals under 30 years old living in Switzerland. A potential solution could be separating different types of plastics at home for collection. The study aims to answer the following research question: To what extent are individuals under 30 in Switzerland aware of the environmental and health impacts of plastic? How do they recycle different types of plastics, and how willing are they to modify their plastic consumption and waste management practices?
:::
# Methods
::: {.justify}
Data was collected via a questionnaire distributed through Google Forms as part of a project for the Research Beyond The Lab course at ETH Zurich. The 21 respondents, all under 30 years old and residing in Zurich, volunteered to complete the questionnaire. The sample was selected to ensure consistency in age, as the study was conducted by an ETH Zurich student. This approach aimed to minimize age-related outliers and provide a focused perspective on the targeted demographic.
:::
# Results
::: {.justify}
This section presents findings related to the second part of the research question. Due to time constraints, the analysis of the first part of the question is omitted. Specifically, this chapter examines the overall plastic recycling rate, the recycling rate per type of plastic, the relationship between the use of single-use plastic bottles and ownership of a reusable bottle, and the willingness to separate different types of plastic at home before collection.
:::
## Plastic recycling rate
::: {.justify}
We start by looking at our respondents. What is the average age, and does everyone recycle plastic? For simplicity, we computed the average age internally and found it to be 24 years old. Regarding the plastic recycling rate, our results were calculated as follows.

```{r, message=FALSE, warning=FALSE, include=TRUE, class.source= "fold-hide"}
#| label: tbl-recycle-rate
#| tbl-cap: "Plastic recycling rate"

# Load required libraries -----------------------------------
library(tidyverse)
library(readxl)
library(janitor)
library(dplyr)
library(knitr)

# Import data -----------------------------------------------
data <- read_csv("../data/processed/plastic_waste_processed.csv")

# Compute age mean ------------------------------------------
age_mean <- mean(data$age, na.rm = TRUE) # 24.25

# Compute plastic recycling rate ----------------------------
recycling_summary <- data |> 
  group_by(rec_plastic) |> 
  summarise(count = n()) |> 
  mutate(percentage = round(count / sum(count) * 100, digits = 1)) |> 
  rbind(c("No", 0, 0)) |> 
  arrange(desc(percentage))

# Create table ----------------------------------------------
recycling_summary |> 
  kable(align = "c")

# Save data underlying the table ----------------------------
write_csv(recycling_summary, 
          here::here("data/final/tbl-recycle-rate.csv"))
```

As shown in @tbl-recycle-rate, the plastic recycling rate is 95.2%, indicating that almost every participant recycles plastic. Moreover, the remaining 4.8% come from 'Sometimes' answers, meaning no one said they do not recycle plastic. This is a promising start. But which type of plastic is recycled the most? Does everyone recycle all types of plastic? To answer this, we create a stacked bar chart displaying the recycling rates for each type of plastic.

```{r, message=FALSE, warning=FALSE, include=TRUE}
#| label: fig-recycled-plastics
#| fig-cap: "Proportion of recycled plastics by type"

# Load required libraries -----------------------------------
library(ggplot2)

# Convert data to long format -------------------------------
data_long <- data |> 
  pivot_longer(cols = rec_PET:rec_PS, names_to = "plastic_type",
               values_to = "rec")

# Calculate the proportions and order data ------------------
data_proportions <- data_long |> 
  group_by(plastic_type, rec) |> 
  summarise(count = n(), .groups = "drop") |> 
  group_by(plastic_type) |> 
  mutate(proportion = count / sum(count)) |> 
  ungroup()

order_data <- data_proportions |> 
  filter(rec == "Yes") |> 
  arrange(desc(proportion))

data_proportions$plastic_type <- factor(data_proportions$plastic_type, 
                                        levels = order_data$plastic_type)

# Create the stacked bar chart ------------------------------
plastic_type_labels <- c("PET", "HDPE", "LDPE", "PP", "PS")

ggplot(data_proportions, aes(x = plastic_type, y = proportion, 
                             fill = factor(rec, 
                                           levels = c("Yes", "No", "Sometimes")))) +
  geom_bar(stat = "identity") +
  scale_y_continuous(labels = scales::label_percent()) +
  scale_x_discrete(labels = plastic_type_labels) +
  scale_fill_manual(values = c("Yes" = "#009999", "No" = "#FF9966", 
                               "Sometimes" = "#FFCCCC"), 
                    labels = c("yes", "no", "sometimes")) +
  labs(
    title = "Plastic recycling rate per type of plastic",
    x = "plastic type",
    y = "proportion",
    fill = "recycle"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(hjust = 1),
        axis.title.x = element_text(margin = margin(t = 10)),
        axis.title.y = element_text(margin = margin(r = 10)), 
        plot.title = element_text(hjust = 0.5, face = "bold", color = "#333333"))

# Save data underlying the figure ---------------------------
write_csv(data_proportions, 
          here::here("data/final/fig-recycled-plastics.csv"))
```

As expected, @fig-recycled-plastics showcases that Polyethylene Terephthalate (PET) - used for water bottles and plastic trays - achieves the highest recycling rate of 100%. Conversely, Low Density Polyethylene (LDPE) - employed in food bags, Polypropylene (PP) - found in margarine tubs and ready-meal trays, and Polystyrene (PS) - utilized for plastic cutlery, are not recycled by the majority of the population, possibly due to a lack of information.
:::
## Relationship between single-use plastic bottle usage and reusable bottle ownership
::: {.justify}
Moving forward, let us delve into the usage of single-use plastic bottles and how it correlates with the ownership of a reusable bottle. To accomplish this, we will employ bar charts.

```{r, message=FALSE, warning=FALSE, include=TRUE}
#| label: fig-bottles
#| fig-cap: "Single-use plastic bottle usage"

# Create summary table --------------------------------------
data <- data |> 
  mutate(own_reusable_bottle = factor(own_reusable_bottle, levels = c("Yes", "No")))

summary_table_bottles <- data |> 
  group_by(use_plastic_bottles, own_reusable_bottle) |> 
  summarise(count = n()) |> 
  ungroup()

order_data_bottles <- summary_table_bottles |> 
  filter(own_reusable_bottle == "Yes") |> 
  arrange(desc(count))

summary_table_bottles$use_plastic_bottles <- 
  factor(summary_table_bottles$use_plastic_bottles, 
         levels = order_data_bottles$use_plastic_bottles)

# Create bar plot
bottles_labels <- c("weekly", "few times a year", "daily", "few times a month", 
                    "never", "monthly")

ggplot(summary_table_bottles, aes(x = use_plastic_bottles, y = count, 
                                  fill = own_reusable_bottle)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_x_discrete(labels = bottles_labels) +
  scale_fill_manual(values = c("Yes" = "#99CCFF", "No" = "#FF9999"), 
                    labels = c("yes", "no")) +
  labs(
    title = "Single-use plastic bottle usage",
    x = "frequency of single-use plastic bottle use",
    y = "count",
    fill = "own reusable bottle"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(hjust = 0.5),
        axis.title.x = element_text(margin = margin(t = 10)),
        axis.title.y = element_text(margin = margin(r = 10)),
        plot.title = element_text(hjust = 0.5, face = "bold", color = "#333333"))

# Save data underlying the figure ---------------------------
write_csv(summary_table_bottles, 
          here::here("data/final/fig-bottles.csv"))
```

As displayed in @fig-bottles, nearly every respondent owns a reusable bottle. However, we did not find a clear link between the frequency of single-use plastic bottle use and the ownership of a reusable bottle. Nonetheless, it is evident that a majority of the population still uses single-use plastic bottles weekly.
:::
## Willingness to separate different types of plastic at home before collection
::: {.justify}
Finally, we examine the willingness to separate plastics at home before collection. Properly collected and sorted household plastics can undergo recycling multiple times [@recycling], highlighting the importance of this practice. We first summarize our data in a table.

```{r, message=FALSE, warning=FALSE, include=TRUE}
#| label: tbl-willingness-investment
#| tbl-cap: "Willingness to separate different types of plastic at home before collection and most significant investment for doing so"

# Create summary table --------------------------------------
data <- data |> 
  mutate(willing_separate = factor(willing_separate, levels = c("Yes", "No")))

summary_table_willingness <- data |> 
  group_by(willing_separate, investment) |> 
  summarise(count = n()) |> 
  arrange(desc(count))

# Display table
summary_table_willingness |> 
  kable(align = "c")

# Save data underlying the table ----------------------------
write_csv(summary_table_willingness, 
          here::here("data/final/tbl-willingness-investment.csv"))
```
From @tbl-willingness-investment, we see that 4 out of 21 respondents are not willing to separate plastics at home before collection, a clear minority. Among those unwilling, space constraints are the most commonly cited reason, while some did not provide a justification. To illustrate the distribution of the most significant investments cited for separating plastics at home, we will create a donut chart.
```{r, message=FALSE, warning=FALSE, include=FALSE}
# Create summary table --------------------------------------
data <- data |> 
  mutate(willing_separate = factor(willing_separate, levels = c("Yes", "No")))

summary_table_willingness <- data |> 
  group_by(investment, willing_separate) |> 
  summarise(count = n()) |> 
  ungroup()

order_data_willingness <- summary_table_willingness |> 
  filter(willing_separate == "Yes") |> 
  arrange(desc(count))

summary_table_willingness$investment <- 
  factor(summary_table_willingness$investment, 
         levels = order_data_willingness$investment)

# Split data for Yes and No --------------------------------
summary_table_yes <- 
  summary_table_willingness |> 
  filter(willing_separate == "Yes") |> 
  select(investment, count) |> 
  rename(Yes = count)

summary_table_no <- 
  summary_table_willingness |> 
  filter(willing_separate == "No") |> 
  select(investment, count) |> 
  rename(No = count)

merged_table <- full_join(summary_table_yes, 
                          summary_table_no, 
                          by = 
                            "investment") |> 
  replace_na(list(Yes = 0, No = 0))

long_table <- merged_table |> 
  pivot_longer(cols = c("Yes", "No"), names_to = "response", values_to = "count")

long_table$investment <- 
  factor(long_table$investment, levels = order_data_willingness$investment)

# Create bar chart -----------------------------------
ggplot(long_table, 
       aes(x = investment, y = count, fill = response)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  scale_y_continuous(breaks = seq(0, 10, 1), limits = c(0, 10)) +
  scale_fill_manual(values = c("Yes" = "#FF9900", "No" = "#990066"), 
                    labels = c("no", "yes")) +
  coord_flip() +
  labs(title = "Willingness to separate vs. most significant investment",
       x = NULL,
       y = "count",
       fill = "willingness") +
  theme_minimal() +
  theme(axis.text.x = element_text(hjust = 0.5),
        axis.title.x = element_text(margin = margin(t = 10)),
        axis.title.y = element_text(margin = margin(r = 10)),
        plot.title = element_text(hjust = 0.5, face = "bold", color = "#333333"))
```

```{r, message=FALSE, warning=FALSE, include=TRUE}
#| label: fig-willingness
#| fig-cap: "Distribution of constraints for separating plastics at home"

# Create summary for donut chart --------------------------
donut_data <- summary_table_willingness |> 
  group_by(investment) |> 
  summarise(count = sum(count)) |> 
  mutate(percentage = count / sum(count))

# Create donut chart --------------------------------------
investment_colors <- c("Space" = "#66C2A5", "Time" = "#FC8D62", "Money" = "#8DA0CB", 
                       "Mental capacity" = "#E78AC3", 
                       "To persue my flatmates to do so aswell" = "#A6D854", 
                       "Where to dispose the trash" = "lightblue", "NA" = "#FFD92F")

ggplot(donut_data, aes(x = 2, y = percentage, fill = investment)) +
  geom_bar(stat = "identity", color = "white") +
  coord_polar(theta = "y", start = 0) +
  xlim(0.5, 2.5) +
  scale_fill_manual(values = investment_colors) +
  labs(title = "Distribution of mentioned 
       constraints related to separating plastics at home",
       fill = "constraint") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", color = "#333333"),
        legend.position = "right")

# Save data underlying the figure ---------------------------
write_csv(donut_data, 
          here::here("data/final/fig-willingness.csv"))
```

In @fig-willingness, we observe that space constraints are identified as the most significant barrier to separating plastics at home, with more than half of the respondents citing it. The second most significant investment is time. These insights highlight the practical challenges individuals face in adopting sustainable waste management practices.
:::
# Conclusions
::: {.justify}
The majors findings of our study can be summarized as follows:

- The overall plastic recycling rate is very high among individuals under 30 years of age living in Zürich. However, Low Density Polyethylene (LDPE) – food bags, Polystyrene (PS) – plastic cutlery, and Polypropylene (PP) – margarine tubs and ready-meal trays are still not recycled by the majority.

- Our analysis did not show any relationship between single-use plastic bottle usage and reusable bottle ownership, but we found that the majority of respondents use single-plastic bottles weekly. This presents an opportunity for easy improvement.

- The willingness to separate plastics at home before collection is strong. However, the lack of space and time poses challenges. Half of the respondents claim that space is the most significant investment needed for this practice. This should be considered if mandatory plastic collection at home is implemented.
:::
# References